---
title: "gold price"
author: "Bhanu Kiran"
date: "12/01/2022"
output:
  word_document: default
  html_document: default
---

# Gold price prediction

## Loading libraries
```{r}
library("readxl")
library(tidyverse)
library(reshape2)
library(caret)
```

Read the dataset
```{r}
gold_df<- read_excel("complete_data.xlsx")
```


```{r}
head(gold_df)
```


```{r}
summary(gold_df)
```


```{r}
str(gold_df)
```

changing the columns names to all lowercase for easy coding
```{r}
cols <- c("date", "sensex", "nifty", "snp", "oil_price", "usdinr", "interest_rate", "gold_demand", "snp_inr", "oil_price_inr", "gold_price")
colnames(gold_df) <- cols
```

```{r}
gold_df$date <- as.character(gold_df$date)
```


we do not need date for correlation, hence making a new dataset
```{r}
sub_gold <- gold_df %>% select(-date)
```


running a correlation on the dataset
```{r}
cormat <- round(cor(sub_gold),2)
head(cormat)
```


```{r}
melted_cormat <- melt(cormat)
head(melted_cormat)
```
Correlation heatmap

```{r}
ggplot(data = melted_cormat, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile()
```

From the above graph everything of the value in light_blue  is positive correlated and everything in navy blue is inversely correlated


## Modelling a linear regression

Exclude the snp and oil_price columns because we take the snp_inr and oil_price_inr as the converted prices
```{r}
mylog <- lm(gold_df$gold_price ~ gold_df$sensex + gold_df$nifty + gold_df$usdinr + gold_df$interest_rate + gold_df$gold_demand + gold_df$snp_inr + gold_df$oil_price_inr)
```


```{r}
summary(mylog)
```

3 statistics to check for goodness of fit
1. p-value => 2.2e-16 which is a very tiny number. This indicates that the values are significant  
2. F-statistic => 798.8 with 2662 degrees of freedom
3. R-squared => 0.7 this is way above 0.5 which indicates a good model

plots of the model

```{r}
plot(mylog)
```

Goodness of fit w.r.t plots
1. Residuals vs Fitted => the variance seems to be equal across the values and the scatter is random
2. Normal qq => slight seems to be mostly normal


```{r}

samples <- gold_df$gold_price %>% createDataPartition(p=0.8, list = FALSE)
train.data <- gold_df[samples,]
test.data <- gold_df[-samples,]
```

```{r}
colnames(gold_df)
```



```{r}
model <- lm(gold_price ~ sensex + nifty + usdinr+ interest_rate+ gold_demand+ snp_inr+ oil_price_inr+ gold_price ,data= train.data)
```

```{r}
summary(model)
```

## Model Interpretation

From our model, before we make our predictions let's have a look at the goodness of fit.

1. R-squared is 0.69 which is ~ .7 and is well above .5 therefor the model is we can say a good model.
2. F-statistic is 601. The higher the better 
3. and p-value very small which means the dependent variable can be explained by independent variables and it is significant

on the predictions below we technically use the formula:

Y= B0 + B1x1 + B2x2 ... + Bnxn + e

where e is the error value which is +/- in nature

according to our predictions whatever value we get is +/- on an average of 7163 units from the actual value


```{r}
predictions <- model %>% predict(test.data)
```

```{r}
RMSE(predictions, test.data$gold_price)
```

```{r}
R2(predictions, test.data$gold_price)
```


```{r}
my_preds <- data.frame(test.data$date, predictions)
```

```{r}
(my_preds)
```

From the table above. We can see that for 8th Jan 2010 the gold price prediction is 62883.17.  
The value in our data set is 51680.7

Our regression formula is Y = prediction +- e  
=> Y = prediction +- 7163  
=> Y = 62883.17 - 7239
=> Y ~ 55720.17

and the accuracy is r-squared value => .69 => 69%  

Therefore the regression formula for predicting the gold prices is:  

Y = 3.631e+03 + (-2.124e+00 x sensex) + (-5.103e-01 x nifty) + (5.408e+02 x usdinr)  
    + (-2.011e+03 x interest rate) + (6.458e+00 x gold demand) + (5.261e-01 x s&p inr) + (1.162e+00 x oil price inr) +- 7239




